\chapter{State of the art} \label{stateoftheart}

In this section the state of the art in conventional and
context-aware recommender systems is presented. 
As a technology recommender systems have been applied 
in many domains, and sometimes they represent the  key
technology for the success of web and mobile applications.

\section{Traditional recommender systems} 
%Tipo de información: Social, Aplicación Social 
%Tags
Some works utilize social information to recommend such
as Manca et.al. \cite{manca2014mining} where the Friend recommender
system is applied for social bookmarking, its goal was to
infer the interest of users from content, selecting the 
information available of the user behavior, and analyzing the resources and the
tags bookmarked for each user. Recommendations are generated
through the mining of user behavior in a tagging system, analyzing the
bookmarks tagged by each user, and the frecuency for each tag used. 
%Keywords
J.Yao et.al. \cite{yao2012product} proposes a new product recommendation
approach for new users based on the implicit relationships between
search keywords and products. The relationships between keywords and
products are represented in a graph and relevance of keywords to
products is derived from attributes of the graph.
%Semantic
The relevance
information is utilized to predict preferences of new users. J.
Golbeck et.al. \cite{golbeck2006filmtrust} presents FilmTrust, a
website that integrates semantic web-based social networks, augmented
with trust, to create prediction of movie recommendations. FilmTrust takes on
the role of a recommender system, forming the core of an algorithm to
predict a rating for movie recommendations. This is an example of
how the Semantic Web, and Semantic Trust networks in particular, can
be exploited to refine the user experience. \\  
%Eso de pros y cons me lo han criticado, me pedían algo más específico
%Este párrafo está bien por que se organiza por problema que atacan las
%propuestas
Traditional recommender techniques have benefits and problems, for
instance, the ability to handle data sparsity and cold-start problems
or considerable ramp-up efforts for knowledge acquisition and
engineering. Establish hybrid systems that combine the strengths of
algorithms and models to overcome some of the shortcomings of
each algorithm.
%De los primeros
A popular example of an hybrid system is called Fab, which was
developed by  Balabanovic et.al. \cite{balabanovic1997fab}. Fab is a
system  for the automatic recognition of emergent issues, relevant to
various  groups of users. It also enables two scaling problems,
pertaining to the rising number of users and documents, to be
addressed.  Claypool et.al. \cite{claypool1999combining}  presents the
P-Tango system that utilizes content-based and collaborative filtering
techniques, it makes a prediction through the weighted average, that
includes content-based prediction and collaborative filtering
prediction. The weights of predictions are determined on a per-user
basis, allowing the system to determine the optimum mixture of
content-based and collaborative recommendation for each user. Pazzani
M. \cite{pazzani1999framework} presents Entree as a hybrid recommender
system that it does not use numeric scores, but rather treats the
output of each recommender (collaborative, content-based and
demographic) as a set of votes, which are then combined in a consensus
scheme. The recommender system includes information such as the
content of the page, ratings of users and demographic data about
users.
%Restaurantes
L. Martinez et al. \cite{martinez2009reja} presents REJA, a hybrid
recommender system that involves collaborative filtering and a
knowledge-based model, that is able to provide recommendations in some
situation for a user; besides, it provides georeferenced information
about the recommended restaurants.
%Turistas
L.Castro et.al. \cite{castro2012prototype} proposes  
a hybrid recommender system for the province of San Juan, Argentina, 
to recommend tourist packages  based on preferences and interest 
of each user, artificial intelligence
techniques are used to filter and customize the information. The
prototype of a recommender system utilizes three techniques to
recommend: demographic, collaborative and content-based. The goal is
to recommend tourist packages that matches the user's profile.
%Hibridos
Others works with hybrid recommender systems are ProfBuilder
  \cite{al1999semantic}, PickAFlick \cite{burke1999integrating}  and
  \cite{tran2000hybrid}, where multiple recommendation
techniques are presented. Usually, a recommendation requires the ranking of
items or the selection of a single best, at this point some
technique must be employed to achieve this. \\ 
Traditional recommender systems (single or hybrid) tend to use
simple user models. For example, user-based collaborative filtering
generally models the user as a vector of item ratings. \\ As additional
observations are made about users' preferences, the user models are
extended, and the users' preferences are used to generate
recommendations. This approach, therefore, ignores the notion of any
specific situation, the fact that users interact with the system
within a particular context and  that preferences of items might 
change in another context. 

\section{Context-aware recommender systems} 
%Contextuales 
Overall, Context factors are able to make a recommender system 
adaptive to the ever changing user's situation. The
context is defined in the domain of the application and the system has
a context model that provides the information for the recommender
system. Some  examples of this types of system are explained
next. \\
Abowd et. al. \cite{abowd1997cyberguide} proposed the
Cyberguide project, which encompass several tour 
guide prototypes for different handheld platforms. 
Cyberguide provided tour guide services to mobile users,
leveraging in the recommendation process both 
\textit{contextual knowledge} about current users and their 
\textit{locations} they have benn in the past. 
Cena et al. \cite{cena2006integrating} presents a tourist guide for particular 
contexts in \textit{intelligent content adaptation}. The UbiquiTO system is a
tourist guide that integrates different forms of context-related
adaptation for: \textit{media device type}, \textit{user characteristics and
preferences}, and the \textit{physical context of the interaction}. UbiquiTO uses
a \textit{rule-based modeling approach} to adapt the content of the provided
recommendation, such as the \textit{amount}, \textit{type of information} and 
\textit{features associated} with each recommendation. 
Bulander et.al \cite{bulander2005comparison} presents the MoMa-system 
that offers \textit{proactive recommendations} using a \textit{post-filtering} 
approach for matching order specifications with offers. 
When creating an order, the client application will automatically fill in both the 
appropriate \textit{physical context} and \textit{profile parameters}, for example, 
for a certain \textit{location} and \textit{weather}, a facility should not be selected
if is too far away from the user's current location and a cold beer should not be
recommended if it is raining. Another feature is related with 
advertisement and offers from suppliers presented in the MoMa-system. These offers are also
formulated according to the catalogue. When the system detects a pair
of context matching order and offer, the end user is notified, in the
preferred manner (for example, SMS, email). At this point, the user
must decide whether to contact the advertiser to accept the offer.
Schifanella et al. \cite{schifanella2008mobhinter} developed
Mob-Hinter, a \textit{context-dependent} distributed model, where a user 
device can directly connect to other mobile devices that are in 
\textit{physical proximity} through ad-hoc connections, hence 
relying on a very limited portion of the users' community and just 
on a subset of all available data (pre-filtering). The relationships 
between users are modeled with a \textit{similarity graph}. Mob-Hinter 
allows a mobile device to identify the affinity network neighbors 
from random ad-hoc communications. The collected information 
is then used to incrementally refine locally
calculated predictions, with no need of interacting with a remote
server or accessing the Internet. Recommendations are computed
using the \textit{available ratings} of the user neighbors.
The PECITAS system \cite{tumas2009personalized}, presented by Thumas, 
offers \textit{location-aware recommendations} for personalized 
\textit{point-to-point paths}. The paths are illustrated by listing 
the various connections that the user must take to reach the 
destination using public transportation and walking. An
interesting aspect of PECITAS is that, although an optimal shortest 
path facility is incorporated, users may be recommended alternative 
routes that pass through several attractions, given that
their specified constraints (e.g. latest arrival time) and travel-related 
preferences (maximum walking time, maximal number of transport
transfers, sightseeing preferences, etc) are satisfied. 
Yu and Chang present LARS  \cite{yu2009personalized} 
which supports personalized tour planning using a \textit{rule-based} 
recommendation process. This system
packages \textit{``where to stay"} and \textit{``where to eat"} 
features together with typical tourist recommendations 
for sightseeing and activities. For
instance, recommended restaurants (selected based on their 
\textit{location}, \textit{menu},  \textit{prices}, \textit{customer rating score}, 
etc.) are integral part of the tour and the time spent for 
lunch/dinner is taken into account to schedule visits to 
attractions or to plan other activities.
Savage et. al. presents the ``I'm feeling LoCo" system \cite{savage2012m}
that proposes a \textit{ubiquitous location­} based recommender algorithm
that focuses in the \textit{user experience} considering 
\textit{user preferences}, \textit{time}, \textit{location} and 
\textit{similarity measures} automatically, using the Foursquare
dataset. The information of the \textit{user's social network},
\textit{transportation} and \textit{phone's sensors} is inferred to provide
a recommendation of places.
Reddy et.al \cite{reddy2006lifetrak} presents the LifeTrack system which 
incorporates sensor information for selecting tracks in a music recomender system.
The songs are represented by tags of terms, that the user assigns in order to link
songs with the \textit{appropriate contexts} in which they should be played.
\textit{User feedback} is incorporated to make a song more or less likely to be
played in a given context. Contextual data considered as relevant to song selection
includes \textit{location}, \textit{time of operation}, \textit{velocity of the user}, 
\textit{weather}, \textit{traffic} and \textit{sound}. \textit{User locations} and 
\textit{velocity} are determined through GPS.
Location information includes tags as \textit{zip code} and whether the
user is \textit{inside or outside} (inferred by the presence or absence of a
GPS signal). The \textit{times of the day} are divided out into configurable
parts of the day (\textit{morning, evening, etc}). The \textit{velocity} is 
abstracted into one of four states: \textit{static, walking, running and driving}.  
Use of accelerometers are planned to enable \textit{indoor velocity}
information. If the user is driving, an RSS feed on traffic information 
is used to typify the state as \textit{calm, moderate or chaotic}. If the user is not
driving, a microphone reading is used for the same purpose.
Additionally, an RSS feed provides a \textit{meteorological condition} (frigid,
cold, temperate, warm or hot).
Ricci et.al. \cite{baltrunas2011incarmusic} uses
context in a music domain, using a model-based paradigm, in this
context-aware recommender system context was defined as a set of
independent contextual factors (independent in order to get a
mathematical model) such as \textit{driving style, road type,
landscape, sleepiness, traffic conditions, mood weather and natural
phenomena} to specifies the relevant context for the music
recommendation. In order to estimate the relevance of selected
contextual factors, users were requested to evaluate music tracks
in different contextual situations for each genre. The prediction
takes in to account this \textit{relevance} to recommend music tracks 
prefered by the user according the \textit{genre} and contexts mentioned.
In the domain of restaurants Chung-Hua et al. \cite{chu2013chinese} proposed a
context-aware recommender system for mobiles using a post-filtering
paradigm, the architecture involves a  client-server model that works
with a request of data from the client side to the server.
Subsequently, taking in account the contextual factors to filter the
appropiate restaurants to recommend. That context-aware recommender
system uses contextual factors such as \textit{location and season}, and
also uses \textit{user preferences} to personalize the recommendations
to a particular context.
Baltrunas et.al. \cite{baltrunas2011context} presents ReRex for tourism, 
a context-aware recommender system, based on a 
\textit{model-based paradigm}, the system recommends and provides 
explanations about why certain Places of Interest(PoI) were recommended. 
The proposed model computes a 
personalized context-dependent rating estimation. Subsequently, in
order to generate the explanation, the system uses
the factor that in the predictive model has the largest positive
effect on the rating prediction. The set of
contextual factors considered in ReRex are: \textit{distance},
\textit{temperature}, \textit{weather}, \textit{season}, \textit{companion}, 
\textit{time day}, \textit{weekday}, \textit{crowdedness}, \textit{familiarity}, 
\textit{mood}, \textit{budget}, \textit{travel length}, \textit{transport} and
\textit{travel goal}. The main issue in ReRex system is the \textit{low user
satisfaction} because the explanations are diffucult to understand,
however, users recognize that the explanation is a very important
component, that influences the system's acceptance. 
Finally, Noguera et. al. \cite{noguera2012mobile} presents a context-aware
recommender system for tourism based in REJA that utilizes the 
\textit{location} through a 3D-GIS system, the application uses 
\textit{progressive downloading} and \textit{rendering of
3D maps} over mobile networks. It is also in charge of tracking the
\textit{user's location} and \textit{speed} based on GPS technology. 
The system utilizes \textit{pre-filtering} and \textit{post-filtering} paradigms. 
Pre-filtering is used to reduce the number of items considered for the
recommendation according to the user's location, and  post-filtering
is applied to re-rank the previous top-N list according to the
physical distance from the user for each recommended restaurant. The
disadvantage in this system is the \textit{lack of user reviews},
because the recommendations are based only in the \textit{location
point} without considering the experience of other diners concerning the
recommended restaurant.\\ To sum up the contextual recommender 
systems mentioned, Table \ref{tab:stateoftheart} describes the main 
contextual factors, the domain of application, the method used for each 
application, and type of devices.
\begin{sidewaystable}[]
  \caption{Comparison of context-aware recommender systems.}
  \label{tab:stateoftheart}
  \bigskip
    \centering\small\setlength\tabcolsep{2pt}
        \hspace*{-1cm}\begin{tabular}{p{3.5cm} p{6cm} p{4cm} p{3cm} p{3cm} }
        %{l l l l l}
           \toprule
             \textbf{Application} &\textbf{Contextual Factor} &\textbf{Domain} &
             \textbf{Paradigm} &\textbf{Device}  \\ \hline
           \midrule
             \textbf{CoMoLE} & \textbf{Time, available time, place, device, level of
            knowledge, learning style.} & \textbf{E-learning} & \textbf{Pre-filtering} 
             & \textbf{Mobiles, PC, laptop.}   \\ \hline 
             \textbf{Moma-System} & \textbf{Location, time.} & 
             \textbf{E-commerce} & \textbf{Post-filtering} & \textbf{PC, laptop.}  \\ 
             \hline
             \textbf{UbiquITO} & \textbf{Season, time, temperature.} & 
             \textbf{Tourism} & \textbf{Post-filtering} & \textbf{Mobiles} \\ \hline
             \textbf{ReRex} & \textbf{Distance of the point of interest,  temperature,
             weather, season, weekend, companion, travel goal, transport.} & 
             \textbf{Tourism} & \textbf{Model-based} & \textbf{Mobiles} \\ \hline
             \textbf{LifeTrack} & \textbf{Location, time, day of the week, traifc noise
             (level), temperature, weather.} & \textbf{Music} & \textbf{ Post-filtering} 
             & \textbf{PC, Mobiles.} \\ \hline
             \textbf{CARS} & \textbf{Location and season.} & \textbf{Restaurants} & 
             \textbf{Post-filtering} & \textbf{PC, laptop.} \\ \hline
             \textbf{InCarMusic} & \textbf{Driving style, road type, landscape, 
             sleepiness, traffic conditions, mood weather and natural phenomena.}
            & \textbf{Music} & \textbf{Model-based} & \textbf{Mobiles} \\ \hline
            \textbf{REJA} & \textbf{Location.} & \textbf{Restaurants} & 
            \textbf{Pre-filtering and Post-filtering} & \textbf{PC, laptop, mobiles.} \\ 
            \hline
            \textbf{CiberGuide} & \textbf{Location, time, weather.} & 
            \textbf{Tourism} & \textbf{Post-filtering} & \textbf{Mobiles} \\ \hline
            \textbf{PECITAS} & \textbf{Location, routes.} & \textbf{Transport} & 
            \textbf{Post-filtering} & \textbf{Mobiles} \\ \hline
            \textbf{LARS} & \textbf{Tourists’ location and time.} & 
            \textbf{Tourist packages} & \textbf{Post-filtering} & \textbf{Mobiles} \\ 
            \hline
            \textbf{I'm feeling LoCo} & \textbf{Location, transportation.} & 
            \textbf{Tourism} & \textbf{Model-based} & \textbf{Mobiles} \\ \hline
            \textbf{MOPSI} & \textbf{Location} & \textbf{Tourism and transport} & 
            \textbf{Post-filtering} & \textbf{Mobiles} \\ \hline
           \bottomrule
        \end{tabular}\hspace*{-1cm}
\end{sidewaystable}








